---
title: "Nashville plots with GWAS summary statistic data"
output: rmarkdown::html_vignette
fig_width: 16
fig_height: 16
vignette: >
  %\VignetteIndexEntry{nashGWAS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE}
#library(ggplot2)
library(nashvillePlot)
```

This document explains how to prepare the data for use with the nashvillePlot package and it's basic usage. The ?functionName command can be used in R to get the documentation of a function.

The pipeline for nashvillePlot is a simple 2 step process.

-   Loading and formatting the input data
-   Generating Plots

## Loading the data

nashvillePlots is most effectivly used to generate two manhattan plots on the same axes with one pointed up and the other pointed down. In this example we will use summary statitics from one gwas and metaXcan results from multiple tissues as the other data set.

The first data set is a single flat file organized like this:

    SNP CHR BP P
    5:85928892 5 85928892 0.4193
    14:3    6082010 14 36082010 0.6898

The second data set will be a number of flat files in a single directory, each organized with the prediXcan output format:

    gene,gene_name,zscore,effect_size,pvalue,var_g,pred_perf_r2,pred_perf_pval,pred_perf_qval,n_snps_used,n_snps_in_cov,n_snps_in_model
    ENSG00000000001,GENE1,-3.27030745841,-2.29584654778,0.00107430648978,0.00115464606277,0.0929513588404,0.00629507630264,0.0191335645147,3,3,3
    ENSG00000000002,GENE2,-3.19327513738,-0.289140191132,0.00140668846561,0.0650699112346,0.0991286399518,0.00471228984588,0.0154733809867,57,58,58

Additionally we need a mapping that provides a correspondence between ensembl id and gene position in the genome. This file is provided in the package

```{r, eval=TRUE}
#[1] "/Users/breyer/Documents/work/giri/Nashville-Plots/dat/"
#metaxcan_results_location <- 
meta.obj <- read_metaXcan_folder(paste0(getwd(), '/../dat/'))
#gwas <- read_gwas_file(file = paste0(getwd(), '/../dat/gwasSumStats.txt')
gwas.obj <- read_gwas_file(file = paste0(getwd(), '/../dat/test.gwas.txt'))
```

## Generating plots

Generate a plot of the whole genome like this, the samp argument is used for purposes of speed as this may take a while to run and the legend is removed for space concerns on this small page

```{r, message=FALSE, eval=TRUE}
options(error=recover)
mygraph <- nashville.plot(data1 = gwas.obj,
                          data2 = meta.obj,
                          data1_direction = 1)
#mygraph <- mygraph + theme(legend.position = "none")
mygraph
```

```{r, echo=FALSE}
knitr::include_graphics("/Users/breyer/Desktop/whole_genome.tiff")
```

Or this:

```{r, message=TR, eval=TRUE}

mygraph <- nashville.plot(data1 = gwas.obj,
                          data2 = meta.obj,
                          data1_direction = -1)
mygraph <- mygraph + theme(legend.position = "none")
mygraph
```

```{r, echo=FALSE}
knitr::include_graphics("/Users/breyer/Desktop/whole_genome2.tiff")
```

Generate a single chromosome plot like this:

```{r, message=FALSE, eval=TRUE}
mygraph <- nashville.plot(data1 = gwas.obj,
                          data2 = meta.obj,
                          chr = 7)
mygraph <- mygraph + theme(legend.position = "none")
mygraph
```

```{r, echo=FALSE}
knitr::include_graphics("/Users/breyer/Desktop/chr7.tiff")
```

or this:

```{r, message=FALSE,eval=TRUE}
mygraph <- nashville.plot(data1 = gwas.obj,
                          data1_direction = -1,
                          chr = 13)
mygraph <- mygraph + theme(legend.position = "none")
mygraph
```

```{r, echo=FALSE}
knitr::include_graphics("/Users/breyer/Desktop/chr13.tiff")
```

Generate a plot around a single gene by using the `zoom_gene` argument or the `zoom_ensg` argument:

```{r, message=FALSE, eval=TRUE}
mygraph <- nashville.plot(data1 = gwas.obj,
                          data2 = meta.obj,
                          chr = 6,
                          zoom_gene = 'GCLC',
                          draw_genes = TRUE)
mygraph <- mygraph + theme(legend.position = "none")
mygraph
```

```{r, echo=FALSE}
knitr::include_graphics("/Users/breyer/Desktop/GCLC_lines.tiff")
```

Or this:

```{r, message=FALSE, eval=TRUE}
options(error=recover)

mygraph <- nashville.plot(data1 = gwas.obj,
                          data2 = meta.obj,
                          chr = 6,
                          data1_direction = -1,
                          zoom_left = 0,
                          zoom_right = 5e6,
                          draw_genes = TRUE)
mygraph <- mygraph + theme(legend.position = "none")
mygraph
```

```{r, echo=FALSE}
knitr::include_graphics("/Users/breyer/Desktop/zoom_manual.pdf")
```
